### 2011-09-17 doc/theses/vitasek/chapter-4-2-experiments-results.tex
Thesis: in progress

Each cell contains the number of times the strategy found optimum in the data set, out of 20 runs. Highlighted are the strategies that performed best on that data set. We see that Strategy 1 and 3 are very similar in performance. We shall nonetheless choose Strategy \textbf{1} as the winner for its simplicity. Now we can tune its parameters.

@@

TODO remind what Strategy 1 is, list all ratios we will be using

TODO we will be gathering for each run the following: what were the parameters, how long did it take and whether it found optimum

@@

@@

\begin{table}
  \caption{Performance of Strategy 1 depending on parameters - excerpt}
  \bigskip
  \label{table-experiments-chained-ihs-tweaking}
  \centering
  \begin{tabular}{c | c | c || c | c | c | c | c | c}
    \heu{RR} & \heu{MUT} & \heu{CX} & \dataset{100-100} & \dataset{100-1000} & \dataset{100-200} & \dataset{OVA1} & \dataset{OVA2} & \dataset{OVA3} \\
    \hline
    0.2	& 0.2	& 0.2	   & 2749.58	& \textbf{8633.76}	  & 3029.82	& 110.82	& 98.68	& 2413.86 \\
    0.5	& 0.05	& 0.05 & 1031.24	& 10324.46	& 1371.08	& \textbf{71.88}	  & 63.86	& 1282.94 \\
    0.5	& 0.05	& 0.2	 & \textbf{919.74}	  & 10978.58	& 1322.02	& 74.46	  & 62.42	& \textbf{1217.00} \\
    0.5	& 0.1	& 0.2	   & 1002.20	& 11030.64	& \textbf{1288.06}	& 85.22	  & 80.96	& 1480.90 \\
    0.5	& 0.2	& 0.1	   & 1674.78	& 9779.82	  & 2007.42	& 108.14	& \textbf{58.56}	& 1827.46 \\
    \end{tabular}
  \bigskip
  First 3 columns are ratios for specified heuristics. Remaining columns are average run times in ms. Bold are the best times in that column.
\end{table}

TODO interpretation - we calculated average runtimes for each combination of ratios and each file. Then we picked the ratio combination that yielded the lowest average runtimes.

RR = 0.5 (strong diversification), MUT = 0.05 (only a few fixed), CX = 0.2 (looking for commonalities among 1/5th of the pool)

TODO compare this to pure GLPK run - for example for \dataset{100-1000}, we got from TODO (90+ on average) seconds to on average TODO seconds (when the 10 second limit is lifted).

@@

TODO we have asked a lot of questions and got the answers, now is the time to summarize, to find some "wisdom".

First of all: if we have the time, it is best to just let the GLPK run.
We will find the optimum in this way.
And for many purposes, this is just fine - we need to infer something about the schema, we do it only once, so it doesn't matter that much how long it takes.

Second: if we don't have enough time, we should do Strategy 1. In under 10 seconds we will have very good results (often the optimum).

% TODO if possible, write down in lessons learned:
% - always allow for interruptions
% - in experiments, iterations should be the outermost loop
% - when representing an object, don't be lazy and create a class for it (don't use pairs, lists with specific data at specific locations, etc)

>>>>

Each cell contains the number of times the strategy found optimum in the data set, out of 20 runs. Highlited are the strategies that performed best on that data set. We see that Strategy 1 and 3 are very similar in performance. We shall nonetheless choose Strategy \textbf{1} as the winner for its simplicity. Now we can tune its parameters.

@@

A short reminder: Scenario 1 consists of \heu{Random} as the CH and the following IHs: \heu{RR} $\rightarrow$ \heu{MUT} $\rightarrow$ \heu{RR} $\rightarrow$ \heu{CX} $\rightarrow$ \heu{RW} $\rightarrow \ldots$

The parameters we can tune in this scenario are the ratios in \heu{RandomRemove} (possibly 2 of them, as there are 2 instances in use), \heu{Mutation} and \heu{Crossover}. We shall not tune the time limits in \heu{Mutation} and \heu{Crossover} and leave them set to 1 second. This presents us with a 3-dimensional space of parameters, where we want to find a combination best suited for our test data sets. We will sample this space by taking a total of 45 configurations of the aforementioned ratios.

@@

This experimental set will consist of 45 ratio combinations * 6 data sets * 50 iterations = 13500 experimental configurations. CH will be \heu{Random} with pool size of 10. IHs will be the ones from Scenario 1, with their ratios set to one of the 45 combinations produced in the following way.
\begin{itemize}
	\item \heu{RandomRemove} ratio will be from $\{0, 0.05, 0.1, 0.2, 0.5\}$
	\item \heu{Mutatio} ratio will be from $\{0.05, 0.1, 0.2\}$
	\item \heu{Crossover} ratio will be from $\{0.05, 0.1, 0.2\}$
\end{itemize}
The process of creating the configurations is captured in Listing \ref{listing-experiment-chained-ihs-tuning}. We will be gathering the following information for each run: what were the parameters, how long did the run take and whether it found optimum.

@@

After averaging the data we get a large result table, an excerpt from which is in Table \ref{table-experiments-chained-ihs-tweaking}. In the left part are the ratio values, in the right part averaged running times for each data set. Highlited are the shortest times for each set. Only rows containing at least one such shortest time are presented.

\begin{table}
  \caption{Performance of Strategy 1 depending on parameters - excerpt}
  \bigskip
  \label{table-experiments-chained-ihs-tweaking}
  \centering
  \begin{tabular}{c | c | c || c | c | c | c | c | c}
    \heu{RR} & \heu{MUT} & \heu{CX} & \dataset{100-100} & \dataset{100-1000} & \dataset{100-200} & \dataset{OVA1} & \dataset{OVA2} & \dataset{OVA3} \\
    \hline
    0.2	& 0.2	& 0.2	   & 2749.58	& \textbf{8633.76}	  & 3029.82	& 110.82	& 98.68	& 2413.86 \\
    0.5	& 0.05	& 0.05 & 1031.24	& 10324.46	& 1371.08	& \textbf{71.88}	  & 63.86	& 1282.94 \\
    0.5	& 0.05	& 0.2	 & \textbf{919.74}	  & 10978.58	& 1322.02	& 74.46	  & 62.42	& \textbf{1217.00} \\
    0.5	& 0.1	& 0.2	   & 1002.20	& 11030.64	& \textbf{1288.06}	& 85.22	  & 80.96	& 1480.90 \\
    0.5	& 0.2	& 0.1	   & 1674.78	& 9779.82	  & 2007.42	& 108.14	& \textbf{58.56}	& 1827.46 \\
    \end{tabular}
\end{table}

It is now necessary to pick one ratio combination as the best one. It is $(RR = 0.5, MUT = 0.05, CX = 0.2)$ for the following reason: it is the best one for sets \dataset{100-100} and \dataset{OVA3} and second best for \dataset{100-200}, \dataset{OVA1} and \dataset{OVA2}. Only the \dataset{100-1000} does not profit from these settings.

Now to interpret ratios in the best combination. \heu{RandomRemove} ratio of 0.5 means that a randomly chosen half of all AMs from every ID set in the pool will be discarded. This amounts to a very strong diversification tendency and keeps the strategy from stalling in local optima. \heu{Mutation} ratio of 0.05 means only around 5\% of AMs in the incumbent solution will be fixed for the next GLPK optimization. \heu{Crossover} ratio of 0.2 means that around 1/5\textsuperscript{th} of ID sets in the pool (randomly chosen) will be scanned for common AMs.

All of the ratios in the best combination are at one end of the range we chose from them. As a future work option it is possible to start moving these ratios even more in their preferred way, possibly removing \heu{Mutation} in the process.

TODO compare this to pure GLPK run - for example for \dataset{100-200}, we got from TODO seconds to on average TODO seconds (when the 10 second limit is lifted).

@@

After asking and answering a lot of questions related to the overall system behavior, parameter effects and various heuristic combinations is now the time to summarize our results and draw conclusions.

The first fact is that if we have the time available, it is best to just let the GLPK run. It will find the optimum eventually, even though this might take minutes or hours to complete. For many purposes, this is just fine - we need to infer something about the schema, we do it only once, so it doesn't matter how long it takes.

Secondly, if we don't have enough time, or have to work in a dynamic environment, we should employ a metaheuristic with a series of improvement heuristics, more specifically Strategy 1. In under 10 seconds we will have very good results, often the optimum.

Overall, it is always good to ignore the simple text data nodes, as it will improve the total search time.

%\section{Lessons Learned}

%This short section will list a few lessons that were learned during the development of \jmodule{IDSetSearch} module for jInfer and experimenting with it.

%The first lesson is that when there is an experiment being run in a few nested loops (see any experiment set construction listing), one of which is \textit{iterations}, this one should be the outermost one. This is because if the experiment somehow fails, TODO

% - always allow for interruptions
% - when representing an object, don't be lazy and create a class for it (don't use pairs, lists with specific data at specific locations, etc)
